Statistical Guarantees for Local Spectral Clustering on Random Neighborhood Graphs

### 2021-11-14 jmlr/revision2/appendix.tex
Copyedits

\subsubsection{Proofs of Lemmas}
In what follows, for a distribution $q$ and vertices $u,w \in V$, we write $q(u,w) := q(u)/d(u) \cdot 1\{(u,w) \in E\}$, and similarly for a collection of dyads $\wt{E} \subseteq V \times V$ we write $q(\wt{E}) := \sum_{(u,w) \in \wt{E}} q(u,w)$. 
\paragraph{Proof of Lemma~\ref{lem:mixing_random_walk}.}
	 We will prove Lemma~\ref{lem:mixing_random_walk} by induction on $t$. In the base case $t = t_0$, observe that $C(K_0;h^{(t_0)}) \cdot \sqrt{\wb{k}} \geq h^{(t_0)}(k) - L_{K_0}(k;h^{(t_0)})$ for all $k \in [K_0, 2m - K_0]$, which implies 
	\begin{equation*}
	L_{K_0}(k;h^{(t_0)}) + C(K_0; h^{(t_0)}) \cdot \sqrt{\wb{k}} \geq h^{(t_0)}(k).
	\end{equation*}

@@

\subsection{Balls, Spherical Caps, and Associated Estimates}
\label{subsec:spherical_caps}
In this section, we derive lower bounds on the volume of the intersection between two balls in $\Rd$, and the volume of a spherical cap. Results of this type are well-known, but since we could not find exactly the statements we desire, for completeness we also supply proofs. We use the notation $B(x,r)$ for a ball of radius $r$ centered at $x \in \Rd$, and $\mathrm{cap}_{r}(h)$ for a spherical cap of height $h$ and radius $r$. Recall that the Lebesgue measure of a spherical cap is
\begin{equation*}
\nu\bigl(\mathrm{cap}_r(h)\bigr) = \frac{1}{2} \nu_d r^d I_{1 - a}\left(\frac{d + 1}{2}; \frac{1}{2}\right),
\end{equation*}
where $a = (r - h)^2/r^2$, and
\begin{equation*}
I_{1 - a}(z,w) = \frac{\Gamma(z + w)}{\Gamma(z) \Gamma(w)} \int_{0}^{1 - a} u^{z - 1} (1 - u)^{w - 1} du,
\end{equation*}
is the cumulative distribution function of a $\mathrm{Beta}(z,w)$ distribution, evaluated at $1 - a$. (Here $\Gamma(\cdot)$ is the gamma function).
\begin{lemma}
	\label{lem:overlap_balls}
	For any $x,y \in \Rd$ and $r > 0$, it holds that
	\begin{equation}
	\label{eqn:overlap_balls_1}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d\biggl(1 - \frac{\|x - y\|}{r} \sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
	For any $x,y \in \Rd$ and $r,\sigma > 0$ such that $\|x - y\| \leq \sigma$, it holds that,
	\begin{equation}
	\label{eqn:overlap_balls_2}
	\nu\bigl(B(x,r) \cap B(y,\sigma)\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
\end{lemma}
\begin{lemma}
	\label{lem:volume_of_spherical_cap}
	For any $0 < h \leq r$, and $a = 1 - (2 r h - h^2)/r^2$,
	\begin{equation*}
	\nu\bigl(\mathrm{cap}_r(h)\bigr) \geq \frac{1}{2}\nu_dr^d\biggl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation*}
\end{lemma}
An immediate implication of~\eqref{eqn:overlap_balls_2} is that for any $x \in \mc{C}_{\lambda,\sigma}$,
\begin{equation}
\label{eqn:uniform_local_conductance}
\nu\bigl(B(x,r) \cap \mc{C}_{\lambda,\sigma}\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
\end{equation}  
\paragraph{Proof of Lemma~\ref{lem:overlap_balls}.}
	First, we prove~\eqref{eqn:overlap_balls_1}. The intersection $B(x,r) \cap B(y,r)$ consists of two symmetric spherical caps, each of height $h = r - \frac{\|x - y\|}{2}$. 
	As a result, by Lemma~\ref{lem:volume_of_spherical_cap} we have
	\begin{equation*}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d \bigl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\bigr)
	\end{equation*}
	where $a = \|x - y\|^2/(4r^2)$, and the claim follows.

@@

\paragraph{Proof of~\eqref{eqn:local_spread_ub}.}
	The statement follows since 
	\begin{equation*}
	d_{\min}(\Pbb) \geq \frac{\pi r^2}{2} \cdot \min_{x \in \mc{X}} f(x) = \frac{\pi r^2}{2} \cdot \frac{\epsilon}{\rho \sigma},
	\end{equation*}
	and
	\begin{equation*}
	\vol_{\Pbb,r}(\mc{X}) \leq d_{\max}(\Pbb) \leq \frac{\pi r^2}{2 \sigma \rho}.
	\end{equation*}
\section{Proof of Major Theorems}
\label{apdx:pf_major_theorems}
We now prove the three major theorems of our paper: Theorem~\ref{thm:volume_ssd_ub} (in Section~\ref{subsec:pf_volume_ssd_ub}), Theorem~\ref{thm:density_cluster_volume_ssd_ub}, and Theorem~\ref{thm:ppr_lb}.  Throughout, we use the notation $\wt{n} = |\mc{C}[X]|$ and $\wt{G}_{n,r} = G_{n,r}\bigl[\mc{C}[X]\bigr]$ as defined above.

@@

\section{Additional results: aPPR and Consistency of PPR}
\label{apdx:appr_misclassification_error}
In this appendix, we prove two additional results regarding PPR remarked upon in our main text. In Section~\ref{subsec:appr_volume_ssd_ub}, we show that clustering using the aPPR vector satisfies an equivalent guarantee to Theorem~\ref{thm:volume_ssd_ub}. In Section~\ref{subsec:consistent_recovery_density_clusters}, we show that the PPR vector can perfectly distinguish two distinct density clusters $\mc{C}_{\lambda},\mc{C}_{\lambda}'$.

@@

\section{Experimental Details}
\label{apdx:experimental_details}
Finally, we detail the settings of our experiments, and include an additional figure.

>>>>

\subsubsection{Proofs of lemmas}
In what follows, for a distribution $q$ and vertices $u,w \in V$, we write $q(u,w) := q(u)/d(u) \cdot 1\{(u,w) \in E\}$, and similarly for a collection of dyads $\wt{E} \subseteq V \times V$ we write $q(\wt{E}) := \sum_{(u,w) \in \wt{E}} q(u,w)$. 
\paragraph{Proof of Lemma~\ref{lem:mixing_random_walk}.}
	 We will prove Lemma~\ref{lem:mixing_random_walk} by induction on $t$. In the base case $t = t_0$, observe that $C(K_0;h^{(t_0)}) \cdot \sqrt{\wb{k}} \geq h^{(t_0)}(k) - L_{K_0}(k;h^{(t_0)})$ for all $k \in [K_0, 2m - K_0]$, which implies 
	\begin{equation*}
	L_{K_0}(k;h^{(t_0)}) + C(K_0; h^{(t_0)}) \cdot \sqrt{\wb{k}} \geq h^{(t_0)}(k).
	\end{equation*}

@@

\subsection{Balls, spherical caps, and associated estimates}
\label{subsec:spherical_caps}
In this section, we derive lower bounds on the volume of the intersection between two balls in $\Rd$, and the volume of a spherical cap. Results of this type are well-known, but since we could not find exactly the statements we desire, for completeness we also supply proofs. We use the notation $B(x,r)$ for a ball of radius $r$ centered at $x \in \Rd$, and $\mathrm{cap}_{r}(h)$ for a spherical cap of height $h$ and radius $r$. Recall that the Lebesgue measure of a spherical cap is
\begin{equation*}
\nu\bigl(\mathrm{cap}_r(h)\bigr) = \frac{1}{2} \nu_d r^d I_{1 - a}\left(\frac{d + 1}{2}; \frac{1}{2}\right),
\end{equation*}
where $a = (r - h)^2/r^2$, and
\begin{equation*}
I_{1 - a}(z,w) = \frac{\Gamma(z + w)}{\Gamma(z) \Gamma(w)} \int_{0}^{1 - a} u^{z - 1} (1 - u)^{w - 1} du,
\end{equation*}
is the cumulative distribution function of a $\mathrm{Beta}(z,w)$ distribution, evaluated at $1 - a$. (Here $\Gamma(\cdot)$ is the gamma function).
\begin{lemma}
	\label{lem:overlap_balls}
	For any $x,y \in \Rd$ and $r > 0$, it holds that
	\begin{equation}
	\label{eqn:overlap_balls_1}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d\biggl(1 - \frac{\|x - y\|}{r} \sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
	For any $x,y \in \Rd$ and $r,\sigma > 0$ such that $\|x - y\| \leq \sigma$, it holds that,
	\begin{equation}
	\label{eqn:overlap_balls_2}
	\nu\bigl(B(x,r) \cap B(y,\sigma)\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
\end{lemma}
\begin{lemma}
	\label{lem:volume_of_spherical_cap}
	For any $0 < h \leq r$, and $a = 1 - (2 r h - h^2)/r^2$,
	\begin{equation*}
	\nu\bigl(\mathrm{cap}_r(h)\bigr) \geq \frac{1}{2}\nu_dr^d\biggl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation*}
\end{lemma}
An immediate implication of~\eqref{eqn:overlap_balls_2} is that for any $x \in \mc{C}_{\lambda,\sigma}$,
\begin{equation}
\label{eqn:uniform_local_conductance}
\nu\bigl(B(x,r) \cap \mc{C}_{\lambda,\sigma}\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
\end{equation}  
\paragraph{Proof of Lemma~\ref{lem:overlap_balls}.}
	First, we prove~\eqref{eqn:overlap_balls_1}. The intersection $B(x,r) \cap B(y,r)$ consists of two symmetric spherical caps, each of height $h = r - \frac{\|x - y\|}{2}$. 
	As a result, by Lemma~\ref{lem:volume_of_spherical_cap} we have
	\begin{equation*}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d \bigl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\bigr)
	\end{equation*}
	where $a = \|x - y\|^2/(4r^2)$, and the claim follows.

@@

\paragraph{Proof of~\eqref{eqn:local_spread_ub}.}
	The statement follows since 
	\begin{equation*}
	d_{\min}(\Pbb) \geq \frac{\pi r^2}{2} \cdot \min_{x \in \mc{X}} f(x) = \frac{\pi r^2}{2} \cdot \frac{\epsilon}{\rho \sigma},
	\end{equation*}
	and
	\begin{equation*}
	\vol_{\Pbb,r}(\mc{X}) \leq d_{\max}(\Pbb) \leq \frac{\pi r^2}{2 \sigma \rho}.
	\end{equation*}

\section{Proof of major theorems}
\label{apdx:pf_major_theorems}
We now prove the three major theorems of our paper: Theorem~\ref{thm:volume_ssd_ub} (in Section~\ref{subsec:pf_volume_ssd_ub}), Theorem~\ref{thm:density_cluster_volume_ssd_ub}, and Theorem~\ref{thm:ppr_lb}.  Throughout, we use the notation $\wt{n} = |\mc{C}[X]|$ and $\wt{G}_{n,r} = G_{n,r}\bigl[\mc{C}[X]\bigr]$ as defined above.

@@

\section{Additional results: aPPR and consistency of PPR}
\label{apdx:appr_misclassification_error}
In this appendix, we prove two additional results regarding PPR remarked upon in our main text. In Section~\ref{subsec:appr_volume_ssd_ub}, we show that clustering using the aPPR vector satisfies an equivalent guarantee to Theorem~\ref{thm:volume_ssd_ub}. In Section~\ref{subsec:consistent_recovery_density_clusters}, we show that the PPR vector can perfectly distinguish two distinct density clusters $\mc{C}_{\lambda},\mc{C}_{\lambda}'$.

@@

\section{Experimental details}
\label{apdx:experimental_details}
Finally, we detail the settings of our experiments, and include an additional figure.
